Case Study: Gene Set Enrichment Analysis (Mouse Data Set)

### 2018-10-31 datasets/GSEA/goseq-human.Rmd
Replace dropbox links with zenodo

The objective of this document is to compare different methods for multiple testing correction in the context of Gene Set Enrichment Analysis (GSEA). In order to get a list of genes, we will use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx data.

>>>>

The objective of this document is to compare different methods for multiple testing correction in the context of Gene Set Enrichment Analysis (GSEA). We use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx individuals. Shortly, samples were downloaded using the *Short Read Archive Toolkit* and mapped to the human reference genome version GRCh38 using *STAR v2.4.2a*. *htseq-count* was used to tabulate the number of uniquely mapping reads for each gene. We used the `DESeq2` package to format the data into a `DESeqDataSet` object.

### 2018-10-31 datasets/GSEA/goseq-mouse.Rmd
Replace dropbox links with zenodo

The objective of this vignette is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA). To do this, we will use mouse data from the paper by [Cabezas-Wallscheid et al. (Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub). The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages. We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.

@@

```{r, message=FALSE}
if (!file.exists(count_file)) {
    download.file("https://www.dropbox.com/s/jq6giqeclxcdypq/gsea-mouse-counts.rds?dl=1",
                  destfile = count_file)
}
dseHSCMPP <- readRDS(count_file)
```

**Why no filtering on gene_biotype == "protein_coding" for this data? AR: I think I did this in the other vignette to keep running times short. This shouldn't change the result since anyway protein coding genes are those with GO categories and genes without GO assignments are ignored.**

>>>>

The objective of this vignette is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA). To do this, we will use mouse data from the paper by [Cabezas-Wallscheid et al. (Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub). The data consist of RNA-seq data from mouse hematopoietic stem cells and multipotent progenitor lineages. The raw fastq data is available through the ArrayExpress database (http://www.ebi.ac.uk/arrayexpress) under accession number E-MTAB-2262. These data were mapped to the mouse reference genome GRCm38 (ENSEMBL release 69) using the Genomic Short-Read Nucleotide Alignment program (version 2012-07-20). We used htseq-count to count the number of reads overlapping with each gene and used the DESeq2 package to format the data as a DESeqDataSet R object.

@@

```{r, message=FALSE}
if (!file.exists(count_file)) {
    download.file("https://zenodo.org/record/1475409/files/gsea-mouse-counts.rds?download=1",
                  destfile = count_file)
}
dseHSCMPP <- readRDS(count_file)
```
